---
title: "CCLE_GDSC_Analysis_Working_21_11_19"
author: "Kieran Redpath"
date: "21 November 2019"
output:
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Shortcuts
Ctrl Shift M = %>% 
Alt - = <- 
Ctrl Alt I = Insert new R chunk
Ctrl Alt M = Git commit

## What exactly are these files?
DepMap-2018q3-celllines.csv: tells you more information about the cell lines
CCLE_mutations.csv: information on what mutations are present in CCLE cell lines
CCLE_RNAseq_genes_counts_20180929.gct: expression data for a bunch of genes in CCLE cell lines

gdsc_codes_1.csv: tells you about drugs (GDSC ID, actual name, etc.)
GDSC_AUC.csv: AUC data for GDSC cell lines, treated with a bunch of different drugs
GDSC_IC50.csv: IC50 data for GDSC cell lines, treated with a bunch of different drugs

# 18_11_19
```{r Load Packages}
setwd("~/Documents/R Data/summer_project")
# Packages
library(limma)
library(edgeR)
library(data.table)
library(magrittr)
library(ggplot2)
library(CePa)
library(dplyr)
library(tidyverse)
```

```{r Data Setup}
# Read Data
depMapCCLEmut <- fread('CCLE_GDSC/Data/CCLE_mutations.csv', sep=',')
dim(depMapCCLEmut)
depMapCCLElines <- fread('CCLE_GDSC/Data/DepMap-2018q3-celllines.csv', sep=',')
dim(depMapCCLElines)
# Check format of cell line info:
head(depMapCCLElines)
# Get cell line barcodes:
bc <- depMapCCLEmut$Tumor_Sample_Barcode
head(bc)
# Match barcodes to Broad IDs
mt <- match(bc, depMapCCLElines$Broad_ID)
sum(is.na(mt))
# Cell lines that don't match to Broad IDs
is.na(mt) %>%  which() %>% bc[.] %>% table()
# Extract disease and gene info
dis <- depMapCCLElines$`Primary Disease`[mt]
gn <- depMapCCLEmut$Hugo_Symbol
```



```{r CDH1}
# Number of CDH1 mutations
sum(gn == "CDH1")
# Lines with CDH1 mutations (ordered by mutation counts)
which(gn == "CDH1") %>% dis[.] %>% table() %>% sort(., decreasing=TRUE) 
# Ignore NSC Lung Cancer for now
diseases = c("Colon Cancer", "Gastric Cancer", "Breast Cancer")
kp <- bc[which(gn == "CDH1" & dis %in% diseases)]
# Colon, breast and gastric cell lines with CDH1 mutations
cdh1Lines <- depMapCCLElines %>% dplyr::filter(., Broad_ID %in% kp)
cdh1Lines %>% 
  dplyr::arrange(., `Primary Disease`) 
col_order_cdh1 <- c("CCLE_Name","Broad_ID","Aliases","COSMIC_ID","Sanger ID","Primary Disease","Subtype Disease","Gender","Source")
cdh1Lines <- cdh1Lines[, col_order_cdh1]
```



```{r Drug Data: Setup}
# Load Data
gdsc_auc <- read.csv('CCLE_GDSC/Data/GDSC_AUC.csv', sep=',')
gdsc_ic50 <- read.csv('CCLE_GDSC/Data/GDSC_IC50.csv', sep=',')
dim(gdsc_auc)
dim(gdsc_ic50)
gdsc_auc[1:5,1:5]
gdsc_ic50[1:5,1:5]
```



```{r Drug Data: Dasatinib}
# GDSC drug codes
drugCodes <- read.csv('CCLE_GDSC/Data/gdsc_codes_1.csv', sep= ',')
head(drugCodes)
# Find Dasatinib (drug_id is "51")
drugCodes %>% dplyr::filter(., drug_name == "Dasatinib")
# Extract Dasatinib data
dasatinib_ic50 <- gdsc_ic50 %>% dplyr::filter(., X=="GDSC:51")
dasatinib_auc <- gdsc_auc %>% dplyr::filter(., X=="GDSC:51")
# Count non-NA values
sum(!is.na(dasatinib_ic50))
sum(!is.na(dasatinib_auc))
# GDSC cell lines with Dasatinib AUC data
names(dasatinib_auc)[!is.na(dasatinib_auc)][-1]
head(dasatinib_ic50)[,1:5]
head(dasatinib_auc)[,1:5]
# Replace names with CCLE names
ic50_names <- match(names(dasatinib_ic50), 
                  gsub("-",".",depMapCCLElines$Broad_ID, fixed=TRUE))
names(dasatinib_ic50) <- depMapCCLElines$CCLE_Name[ic50_names]
head(dasatinib_ic50)[,1:5]

auc_names <- match(names(dasatinib_auc), 
                  gsub("-",".",depMapCCLElines$Broad_ID, fixed=TRUE))
names(dasatinib_auc) <- depMapCCLElines$CCLE_Name[auc_names]
head(dasatinib_auc)[,1:5]
```



```{r CCLE Expression Data Setup/Test}
# Load data, convert to table
ccleExpData <- read.gct('CCLE_GDSC/Data/CCLE_RNAseq_genes_counts_20180929.gct')
dim(ccleExpData)
# Data are read counts and ensembl gene IDs
ccleExpData[1:5,1:5]
# Log the data (add 0.5 to avoid log(0))
logExpDat <- log(ccleExpData + 0.5)
# Density plots to assess consistency-> data are consistent with one another -> will normalise this next
plot(density(logExpDat[,1]))
for(i in 2:ncol(logExpDat)) lines(density(logExpDat[,i]))
```



```{r CCLE Expression Data Normalised/Unfiltered}
# Normalise via limma::voom (see page 70 of Limma usersguide)
# Note - I'm not filtering the genes
dge <- DGEList(counts=ccleExpData)
dge <- calcNormFactors(dge)

v <- voom(dge, plot=TRUE)

# Extract the normalised expression data
# Note that the voom procedure produces logged data
expDat <- v$E

# Plot the post-normalised data -> the important part
# Note there is now less variation in the "bump" above 5 - instead the 
# variation has been "moved" to the very low abundance genes (below -5).
plot(density(expDat[,1]))
for(i in 2:ncol(expDat)) lines(density(expDat[,i]))

expDat[1:5,1:5]
```



## Now you can start having a play with the expression data in combination with the Dasatinib AUC and IC50 information. Can also add in CDH1 mutation data as needed.

# 19_11_19
```{r Extract Tissue Info for IC50, AUC, and expDat}}
#Create a vector containing the shared sample names (the intersect of identical data)
commonSamples <- intersect(names(dasatinib_ic50),colnames(expDat))
head(commonSamples)

# Figure out which IC50 observations are within commonSamples (match)
ic50_match <- match(commonSamples, names(dasatinib_ic50))
# Interpret this (e.g. the 18th entry in dasatinib_ic50 matches the first entry in commonSamples)
head(ic50_match)
# Figure out which expDat observations are within commonSamples (match)
expDat_match <- match(commonSamples, colnames(expDat))
head(expDat_match)

# Create a new IC50 variable containing samples from commonSamples (na.omit) and sorted in the same order (also due to na.omit)
dasatinib_ic50_sort <- dasatinib_ic50[na.omit(ic50_match)]
# Create a new expDat variable containing samples from commonSamples (na.omit) and sorted in the same order (also due to na.omit)
expDat_sort <- expDat[ , na.omit(expDat_match)]

# strsplit takes each "names" data, splitting it at each "_", creating a list.
# lapply takes each list entry from this, dropping the first value (in this case, cell-line), then pastes the rest back together with "_".
# unlist converts the data from a list back to a vector.
# Define "tissue_sort"
tissue_sort <- strsplit(names(dasatinib_ic50_sort), "_") %>% 
  lapply(., function(x) paste0(x[-1], collapse='_')) %>%  
  unlist()
# Check numbers for each tissue type
table(tissue_sort)

# Create a vector containing the shared sample names (the intersect of identical data)
commonSamples <- intersect(names(dasatinib_auc),colnames(expDat))
head(commonSamples)
# Figure out which IC50 observations are within commonSamples (match)
auc_match <- match(commonSamples, names(dasatinib_auc))
# Interpret this 
head(auc_match)
# Figure out which expDat observations are within commonSamples (match)
expDat_auc_match <- match(commonSamples, colnames(expDat))
head(expDat_match)
# Create a new auc variable containing samples from commonSamples (na.omit) and sorted in the same order (also due to na.omit)
dasatinib_auc_sort <- dasatinib_auc[na.omit(auc_match)]
# Create a new expDat variable containing samples from commonSamples (na.omit) and sorted in the same order (also due to na.omit)
expDat_auc_sort <- expDat[ , na.omit(expDat_auc_match)]

# Note that the order is now the same
expDat_auc_sort[1:5,1:5]
dasatinib_auc_sort[1:5]
dasatinib_ic50_sort[1:5]
```



# 20_11_19

## CCLE_dasat_BCCCGC is *all* BC, CC, and GC lines with GDSC data
```{r Find Information on Dasatinib Treatment for CDH1 Mutant Breast, Colon and Gastric Cancer Lines}
# Create the filtered file, removing cell line info. Leaves you with IC50 for BC, CC, GC
tissue_sort_BCCCGC <- strsplit(names(dasatinib_ic50_sort), "_") %>% 
  lapply(., function(x) paste0(x[-1], collapse='_')) %>%  
  unlist()
dasat_ic50_BCCCGC <- dasatinib_ic50_sort[tissue_sort == "STOMACH" | tissue_sort == "BREAST" | tissue_sort == "LARGE_INTESTINE"]
# Create the filtered file, removing cell line info. Leaves you with AUC for BC, CC, GC
tissue_sort_BCCCGC2 <- strsplit(names(dasatinib_auc_sort), "_") %>% 
  lapply(., function(x) paste0(x[-1], collapse='_')) %>%  
  unlist()
dasat_auc_BCCCGC <- dasatinib_auc_sort[tissue_sort_BCCCGC2 == "STOMACH" | tissue_sort_BCCCGC2 == "BREAST" | tissue_sort_BCCCGC2 == "LARGE_INTESTINE"]
# Combine these two files
dasat_data_BCCCGC <- rbind(dasat_auc_BCCCGC, dasat_ic50_BCCCGC)
rownames(dasat_data_BCCCGC) <- c("AUC","IC50")
dasat_data_BCCCGC <- t(dasat_data_BCCCGC)

# Create a file with cell line info to combine with "dasat_data_BCCCGC"
CCLElinesBCCCGC <- depMapCCLElines %>%
  filter(`Primary Disease`=="Gastric Cancer"|`Primary Disease`=="Breast Cancer" |`Primary Disease`=="Colon Cancer")
# Rename CCLElinesBCCCGC rows
l <- as.list(CCLElinesBCCCGC$CCLE_Name)
rownames(CCLElinesBCCCGC) <- l
CCLElinesBCCCGC <- select (CCLElinesBCCCGC, -c(CCLE_Name))

#Combine dasatinib data and CCLElinesBCCCGC
CCLE_dasat_BCCCGC <- merge(CCLElinesBCCCGC,dasat_data_BCCCGC,by=0, all=T)
names(CCLE_dasat_BCCCGC)[names(CCLE_dasat_BCCCGC) == 'Row.names'] <- 'CCLE_Name'
View(CCLE_dasat_BCCCGC)
```



# 21_11_19

## Create a file with CDH1 non-mutants
```{r CDH1 Non-mutants}
# Number of lines without CDH1 mutations
sum(gn != "CDH1")
# Lines without CDH1 mutations (ordered by mutation counts)
which(gn != "CDH1") %>% dis[.] %>% table() %>% sort(., descending = TRUE)
# Ignore NSC Lung Cancer for now
diseases = c("Colon Cancer", "Gastric Cancer", "Breast Cancer")
kk <- bc[which(gn != "CDH1" & dis %in% diseases)]
# Colon, breast and gastric cell lines without CDH1 mutations
Noncdh1Lines <- depMapCCLElines %>% dplyr::filter(., Broad_ID %in% kk)
Noncdh1Lines %>% 
  dplyr::arrange(., `Primary Disease`)
col_order_Noncdh1 <- c("CCLE_Name","Broad_ID","Aliases","COSMIC_ID","Sanger ID","Primary Disease","Subtype Disease","Gender","Source")
Noncdh1Lines <- Noncdh1Lines[, col_order_Noncdh1]

# Open the new data
View(Noncdh1Lines)
View(cdh1Lines)
```



# 25_11_19

## Add CDH1 mutation data to the existing table, then create a boxplot with CDH1 mutants and non-CDH1 mutants next to each other, telling you about IC50 values.
```{r IC50 and AUC for Mutants vs Non-mutants Test, echo=FALSE, include=FALSE}
# Add CDH1_Mutations column to CCLE_dasat_BCCCGC
CCLE_dasat_BCCCGC$CDH1_Mutations=0
# Add AUC, IC50, and CDH1_Mutations columns to CCLE_dasat_BCCCGC
cdh1Lines$AUC=NA
cdh1Lines$IC50=NA
cdh1Lines$CDH1_Mutations=1
# Use match to combine the two
Boxplot_data <- match(cdh1Lines$CCLE_Name, CCLE_dasat_BCCCGC$CCLE_Name)
CCLE_dasat_BCCCGC$CDH1_Mutations[Boxplot_data] <-  1

# The actual boxplot
ggplot(data = CCLE_dasat_BCCCGC, mapping =aes(y=AUC)) +
  geom_boxplot()

ggplot(data = CCLE_dasat_BCCCGC, mapping =aes(group=CDH1_Mutations, y=AUC)) +
  geom_boxplot()
```

```{r AUC and IC50 Plots}
CCLE_dasat_BCCCGC$CDH1_Mutations=FALSE
# Add AUC, IC50, and CDH1_Mutations columns to CCLE_dasat_BCCCGC
cdh1Lines$AUC=NA
cdh1Lines$IC50=NA
cdh1Lines$CDH1_Mutations=TRUE
# Use match to combine the two
Boxplot_data <- match(cdh1Lines$CCLE_Name, CCLE_dasat_BCCCGC$CCLE_Name)
CCLE_dasat_BCCCGC$CDH1_Mutations[Boxplot_data] <-  TRUE

ggplot(data = CCLE_dasat_BCCCGC, mapping =aes(x=CDH1_Mutations, y=AUC)) +
  geom_boxplot()

ggplot(data = CCLE_dasat_BCCCGC, mapping =aes(x=CDH1_Mutations, y=IC50)) +
  geom_boxplot()
```
